Standard Candle Cosmology With Incomplete Spectroscopic Classification

### 2015-11-12 doc/article.tex

Type~Ia supernovae (SNe~Ia) are an example of a standard candle (more precisely a standardizable
candle), whose redshift-magnitude relation provides an accurate expansion history
used to measure the Hubble Constant
\citep{2001ApJ...553...47F} and detect the accelerated expansion of the Universe
\citep{1998AJ....116.1009R, 1999ApJ...517..565P}.  SN~Ia measurements continue
to improve 
\citep{2014A&A...568A..22B} and play a critical role in probing the physics
responsible for the acceleration \citep{2013PhR...530...87W}.
Looking at the present and toward the future, SN~Ia cosmology is
a component of the Dark Energy Survey\footnote{\url{http://www.darkenergysurvey.org}} (DES),
and a science driver in the design of the
Large Synoptic Survey Telescope\footnote{\url{http://www.lsst.org}} (LSST)
and the Wide-Field InfrarRed Survey Telescope
\citep{2015arXiv150303757S}.

@@

The Probabilistic Graphical Model is shown in Figure~\ref{pgm:fig}. The underlying behavior of the universe is described by the following elements:
\begin{description}
\item [Cosmology] Luminosity distance for the cosmological model being considered,
e.g.\ a $wCDM$ universe parameterized by $\Omega_M$, $\Omega_w$, $w_0$, $w_a$.
\item[SNe~Ia, Non-Ia Populations] Time-dependent luminosity for all transient populations
that may enter
the analysis.  The SALT2 model is a candidate for SNe~Ia, with parameters $\alpha$, $\beta$.  The PSNID treatment of non-Ia's can be made into a model for the other
transients or Natahsa Karpenka is working on a model of non-Type~II core collapse SNe.
\item [Relative Rates] Relative rates of all the transients populations and
subpopulations (e.g. SALT2 $x_1$, $c$) under consideration.  The
model could of been extended to measure absolute rates, but that is unnecessary if
the cosmological parameters are the goal.
\end{description}
Each transient is described by the following elements:
\begin{description}
\item[Host Galaxy] The host galaxy, or more generally for the case of hostless supernovae,
the environment from which the transient arises.  The host is described by its redshift
and other parameters that may be of interest such as mass,
specific star formation rate, metallicity.
\item[Type Subtype] Specification of the the transient: the population to which it belongs
and its internal parameters.  For example, an object could be SN~Ia with specific
values for its SALT2 parameters $c$, $x_1$, $t_0$ and its random dispersion
realization (both absolute magnitude and from the error snake). 
This is dependent on the relative rates and
its host galaxy including its redshift.
\item[Distance] Distance modulus or equivalently the luminosity distance of the host galaxy,
which in our model corresponds to the cosmological redshift of the transient.  The value
is fixed based on the cosmology and redshift.
\end{description}
The individual photometric measurements of a transient are described by the following
elements:
\begin{description}
\item[Luminosity]  Luminosity as a function of wavelength
corresponding to the date of observation.   This depends on the population, type,
and subtype.  Given that searching for correlations between  host properties and Hubble
residuals is a good probe of systematics,
a direction connection between host galaxy and luminosity is made.
\item[Flux] Flux of the transient incident to earth at the date of observation.  It is 
fixed based on the luminosity, redshift, and distance.
\end{description}
The observatory is described by the following element:
\begin{description}
\item[Throughput] The transmission function experienced by the transient during
the observation.  It depends on band, position on the focal plane, and date of observation.
\end{description}
The data (subscripted by ``o'' for observation)  are:
\begin{description}
\item[$\text{Photometry}_o$] Photometric flux measurement of the transient.
The expected measurement
is the incident flux integrated over the throughput.
\item[$\text{Type Subtype}_o$] Spectroscopic
measurements of the intrinsic transient type.  It is dependent on the underlying
type.
\item[$\text{Redshift}_o$] Measurement of the host redshift.  The redshift can be
spectroscopic or photometric.  It is dependent on the host redshift, and
redshifts of any other galaxies that may be confused as the host (not shown explicitly
in the PGM).
\item[$\text{Host Properties}_o$] Measurement of the host properties such
as mass, sSFR, metallicity. It is dependent on the host galaxy, and
any other galaxies that may be confused as the host.
\end{description}

@@

\subsection{Detection Criteria (Malmquist Bias)}
The photometric selection criteria are expected to be significantly more rigorous that
the detection criteria.  Nevertheless, the detection efficiencies need to be characterized.
Fakes already provide a good measure of this.  To first order, the efficiency depends
on $S/N$, but a secondary dependence on the structure of background galaxies.
The detection efficiency is more complicated than the step function
shown in Eqn.~\ref{eff:eqn}.

@@

\subsection{Per-Exposure Thresholds}
To understand the effect of sample-selection criteria and detection efficiency
on a specific object, the
properties of each photometric point is needed.  Although the photometric criteria
have not been formalized, a $S/N$ threshold is a possible statistic to be used.
As such, a clean deliverable would be:
\begin{itemize}
\item Determination of flux thresholds for each photometric point of each candidate
given
a $S/N$
\begin{equation}
threshold(point, S/N).
\end{equation}
If we work in the sky-noise dominated regime, I believe this information is already
calculated by DM.
\end{itemize}

>>>>

Type~Ia supernovae (SNe~Ia) are an example of a standard candle (more precisely a standardizable
candle), whose redshift-magnitude relation provides an accurate expansion history
used to measure the Hubble Constant
\citep{2001ApJ...553...47F} and detect the accelerated expansion of the Universe
\citep{1998AJ....116.1009R, 1999ApJ...517..565P}.  SN~Ia measurements continue
to improve 
\citep{2014A&A...568A..22B} and play a critical role in probing the physics
responsible for the acceleration \citep{2013PhR...530...87W}.
Looking at the present and toward the future, SN~Ia cosmology is
a component of the Dark Energy Survey\footnote{\url{http://www.darkenergysurvey.org}} (DES),
and a science driver in the design of the
Large Synoptic Survey Telescope\footnote{\url{http://www.lsst.org}} (LSST)
and the Wide-Field InfraRed Survey Telescope
\citep{2015arXiv150303757S}.

@@

The Probabilistic Graphical Model is shown in Figure~\ref{pgm:fig}. The underlying behavior of the universe is described by the following elements:
\begin{description}
\item [Cosmology] Luminosity distance for the cosmological model being considered,
e.g.\ a $wCDM$ universe parameterized by $\Omega_M$, $\Omega_w$, $w_0$, $w_a$.
\item[SNe~Ia, Non-Ia Populations] Time-dependent luminosity for all transient populations
that may enter
the analysis.  The SALT2 model is a candidate for SNe~Ia, with parameters $\alpha$, $\beta$.  The PSNID treatment of non-Ia's can be made into a model for the other
transients or Natahsa Karpenka is working on a model of non-Type~II core collapse SNe.
\item [Relative Rates] Relative rates of all the transients populations and
subpopulations (e.g. SALT2 $x_1$, $c$) under consideration.  The
model could of been extended to measure absolute rates, but that is unnecessary if
the cosmological parameters are the goal.
\end{description}
Each transient is described by the following elements:
\begin{description}
\item[Host Galaxy] The host galaxy, or more generally for the case of hostless supernovae,
the environment from which the transient arises.  The host is described by its redshift
and other parameters that may be of interest such as mass,
specific star formation rate, metallicity.
\item[Type Subtype] Specification of the the transient: the population to which it belongs
and its internal parameters.  For example, an object could be SN~Ia with specific
values for its SALT2 parameters $c$, $x_1$, $t_0$ and its random dispersion
realization (both absolute magnitude and from the error snake). 
This is dependent on the relative rates and
its host galaxy including its redshift.
\item[Distance] Distance modulus or equivalently the luminosity distance of the host galaxy,
which in our model corresponds to the cosmological redshift of the transient.  The value
is fixed based on the cosmology and redshift.
\end{description}
The individual photometric measurements of a transient are described by the following
elements:
\begin{description}
\item[Luminosity]  Luminosity as a function of wavelength
corresponding to the date of observation.   This depends on the population, type,
and subtype.  Given that searching for correlations between  host properties and Hubble
residuals is a good probe of systematics,
a direct connection between host galaxy and luminosity is made.
\item[Flux] Flux of the transient incident to earth at the date of observation.  It is 
fixed based on the luminosity, redshift, and distance.
\end{description}
The observatory is described by the following elements:
\begin{description}
\item[Throughput] The transmission function experienced by the transient during
the observation.  It depends on band, position on the focal plane, and date of observation.
\end{description}
The data (subscripted by ``o'' for observation)  are:
\begin{description}
\item[$\text{Photometry}_o$] Photometric flux measurement of the transient.
The expected measurement
is the incident flux integrated over the throughput.
\item[$\text{Type Subtype}_o$] Spectroscopic
measurements of the intrinsic transient type.  It is dependent on the underlying
type.
\item[$\text{Redshift}_o$] Measurement of the host redshift.  The redshift can be
spectroscopic or photometric.  It is dependent on the host redshift, and
redshifts of any other galaxies that may be confused as the host (not shown explicitly
in the PGM).
\item[$\text{Host Properties}_o$] Measurement of the host properties such
as mass, sSFR, metallicity. It is dependent on the host galaxy, and
any other galaxies that may be confused as the host.
\end{description}

@@

\subsection{Detection Criteria (Malmquist Bias)}
The photometric selection criteria are expected to be significantly more rigorous than
the detection criteria.  Nevertheless, the detection efficiencies need to be characterized.
Fakes already provide a good measure of this.  To first order, the efficiency depends
on $S/N$, but a secondary dependence on the structure of background galaxies.
The detection efficiency is more complicated than the step function
shown in Eqn.~\ref{eff:eqn}.

@@

\subsection{Per-Exposure Thresholds}
To understand the effect of sample-selection criteria and detection efficiency
on a specific object, the
properties of each photometric point are needed.  Although the photometric criteria
have not been formalized, a $S/N$ threshold is a possible statistic to be used.
As such, a clean deliverable would be:
\begin{itemize}
\item Determination of flux thresholds for each photometric point of each candidate
given
a $S/N$
\begin{equation}
threshold(point, S/N).
\end{equation}
If we work in the sky-noise dominated regime, I believe this information is already
calculated by DM.
\end{itemize}
